Mixing Time Estimation in Reversible Markov Chains from a Single Sample Path

### 2015-06-09 arxiv-v1/intro.tex
orphans

\paragraph{Main results.}
We develop the first procedure for constructing non-trivial
and fully empirical confidence intervals for Markov mixing time.
Consider a reversible ergodic Markov chain on $d$ states with absolute
spectral gap $\gap$ and stationary distribution minorized by $\pimin$. 
As is well-known \citep[Theorems~12.3 and~12.4]{LePeWi08},
\begin{equation}
  \label{eq:mixing-time-bound}
  \Parens{\mfrac1{\gap} - 1} \ln2
  \ \leq \ \tmix
  \ \leq \ \mfrac1{\gap} \ln \mfrac4{\pimin}
  ,
\end{equation}
and hence it suffices to estimate $\gap$ and $\pimin$. 
Our main results are summarized as follows.
\begin{enumerate}
  \item
    In \cref{sec:rates-lower}, we show that in some problems 
    $n = \Omega\parens{ (d\log d)/\gap + 1/\pimin}$ observations are necessary for any procedure
    to guarantee constant multiplicative accuracy in estimating $\gap$
    (\cref{thm:lb-pimin,thm:lb-gap}).
    Essentially, in some problems \emph{every} state may need to be visited about
    $\log(d)/\gap$ times, on average, before an accurate estimate of
    the mixing time can be provided, regardless of the estimation
    procedure used.

>>>>

\paragraph{Main results.}
We develop the first procedure for constructing non-trivial
and fully empirical confidence intervals for Markov mixing time.
Consider a reversible ergodic Markov chain on $d$ states with absolute
spectral gap $\gap$ and stationary distribution minorized by $\pimin$. 
As is well-known \citep[Theorems~12.3 and~12.4]{LePeWi08},
\begin{equation}
  \label{eq:mixing-time-bound}
  \Parens{\mfrac1{\gap} - 1} \ln2
  \ \leq \ \tmix
  \ \leq \ \mfrac1{\gap} \ln \mfrac4{\pimin}
  ,
\end{equation}
and hence it suffices to estimate $\gap$ and $\pimin$. 
Our main results are summarized as follows.
\begin{enumerate}
  \item
    In \cref{sec:rates-lower}, we show that in some problems 
    $n = \Omega\parens{ (d\log d)/\gap + 1/\pimin}$ observations are necessary for any procedure
    to guarantee constant multiplicative accuracy in estimating $\gap$
    (\cref{thm:lb-pimin,thm:lb-gap}).
    Essentially, in some problems \emph{every} state may need to be visited about
    $\log(d)/\gap$ times, on average, before an accurate estimate of
    the mixing time can be provided, regardless of the actual estimation
    procedure used.
